\section{Computation Graphs \& Backpropagation}
\slideref{Session 4}

These notes accompany the fourth lecture. The central question left open at the end of Session~3 was: how do we compute $\nabla_{\pvec} \loss(\pvec)$ efficiently for a deep network? The answer --- backpropagation --- is the algorithmic backbone of all modern deep learning. We develop it here through the framework of computation graphs.

% - - - - - - - - - - - - - - - - - --
\subsection{Forward and Backward Pass}
\slideref{Session 4: Forward and Backward Pass}

Training a deep network requires two passes through the network for each update step. The \textbf{forward pass} evaluates the network layer by layer from input to output, computing the prediction $\yhat$ and then the loss $\loss$:
\begin{equation}
\vx = \vh^{(0)} \xrightarrow{\thetamat{1}} \vh^{(1)} \xrightarrow{\thetamat{2}} \cdots \xrightarrow{\thetamat{L}} \vh^{(L)} = \yhat \xrightarrow{\ell} \loss.
\end{equation}
The \textbf{backward pass} propagates gradients in the reverse direction, from the loss back through each layer to the parameters:
\begin{equation}
\frac{\partial \loss}{\partial \pvec} \xleftarrow{\thetamat{1}} \frac{\partial \loss}{\partial \vh^{(1)}} \xleftarrow{\thetamat{2}} \cdots \xleftarrow{\thetamat{L}} \frac{\partial \loss}{\partial \vh^{(L)}} \xleftarrow{\ell} \loss.
\end{equation}
The \textbf{gradient} $\nabla_{\pvec}\,\loss(\pvec) = \frac{\partial \loss}{\partial \pvec}$ collects one partial derivative per parameter. Backpropagation is the algorithm that computes this backward pass efficiently. It is important to note that backpropagation \emph{computes gradients} --- it does not perform any optimisation. The gradient descent update step is separate.

% - - - - - - - - - - - - - - - - - --
\subsection{Computation Graphs}
\slideref{Session 4: Computation Graphs --- Motivation, Example, Flexible Level of Detail, Flexible Operations, Scalar vs Vector}

A \textbf{computation graph} is a directed acyclic graph in which each node represents an elementary operation and each edge represents the flow of data between operations. Decomposing a complex function into a computation graph makes gradient computation tractable for two reasons. First, each atomic operation has a simple, easy-to-compute local derivative. Second, intermediate values computed during the forward pass can be stored and reused during the backward pass, avoiding redundant computation.

A computation graph has three types of nodes:
\begin{itemize}
    \item \textbf{Input nodes} --- hold fixed data values ($x$, $y$). They store a value but have no gradient.
    \item \textbf{Parameter nodes} --- hold learnable parameters ($\pvec$). They store a value and accumulate a gradient.
    \item \textbf{Compute nodes} --- implement elementary operations. They store a value, accumulate a gradient, and know how to compute both their forward function and their local gradient.
\end{itemize}

\subsubsection{Example: Linear Regression}

Consider linear regression with MSE loss: $\ell(x, y, \pvec) = (\theta_0 + \theta_1 x - y)^2$. We decompose this into four atomic steps:
\begin{align}
(1)\ & h = \theta_1 x \\
(2)\ & \yhat = \theta_0 + h \\
(3)\ & z = \yhat - y \\
(4)\ & \ell = z^2.
\end{align}
Each step is a simple operation --- multiplication, addition, subtraction, squaring --- with a derivative that can be written down immediately.

\subsubsection{Flexible Level of Detail}

The same function can be decomposed at different levels of granularity. The four-step decomposition above is fine-grained; we could also use a coarser two-step decomposition:
\begin{align}
(1)\ & \yhat = \theta_0 + \theta_1 x \\
(2)\ & \ell = (\yhat - y)^2,
\end{align}
or a vectorised one-step version $\yhat = \pvec^\top \vx$. The choice of granularity is a design decision: finer decompositions make each local gradient simpler, while coarser ones reduce the number of nodes and may be more efficient if the compound derivative is known analytically.

\subsubsection{Flexible Operations}

The same graph structure can represent different models by changing the node functions. A two-node graph with an intermediate prediction node $\yhat$ and a loss node $\ell$ covers both linear regression with MSE and logistic regression with binary cross-entropy --- only the functions at each node differ. This modularity is the foundation of modern deep learning frameworks, which implement a library of differentiable operations that can be composed freely.

\subsubsection{Scalar vs Vector Notation}

Computation graphs can be drawn at the level of scalar operations --- one node per neuron, one edge per weight --- or at the level of vector operations --- one node per layer. Scalar graphs make the gradient flow explicit; vector graphs are more compact and correspond directly to the matrix operations implemented in practice. The two representations compute the same quantities; the vector form simply groups many scalar operations into a single matrix node.

% - - - - - - - - - - - - - - - - - --
\subsection{Backpropagation}
\slideref{Session 4: Two Key Rules, Local vs Global Gradients}

Backpropagation relies on two rules from calculus.

\textbf{Sum rule.} The gradient of an average is the average of the gradients:
\begin{equation}
\nabla_{\pvec}\, \loss(\pvec)
= \nabla_{\pvec} \frac{1}{\nsamples}\sum_{i=1}^{\nsamples} \ell(\vx^{(i)}, y^{(i)}, \pvec)
= \frac{1}{\nsamples}\sum_{i=1}^{\nsamples} \nabla_{\pvec}\, \ell(\vx^{(i)}, y^{(i)}, \pvec).
\end{equation}
This means we can compute the gradient for a single training sample and average the results --- we never need to differentiate through the sum directly.

\textbf{Chain rule.} For a composed function $f(g(x))$:
\begin{equation}
\frac{d}{dx} f(g(x)) = \frac{df}{dg}\frac{dg}{dx}.
\end{equation}
When a variable $x$ feeds into multiple downstream nodes $g_1(x), \ldots, g_m(x)$, the chain rule generalises to:
\begin{equation}
\frac{d}{dx} f(g_1(x), \ldots, g_m(x)) = \sum_{i=1}^{m} \frac{\partial f}{\partial g_i}\frac{dg_i}{dx}.
\end{equation}
The sum arises because $x$ influences $f$ through multiple paths; the total effect is the sum of contributions from each path.

\subsubsection{Local and Global Gradients}

At each node in the computation graph we distinguish two quantities:

\begin{itemize}
    \item The \textbf{local gradient} $\frac{\partial\, \text{out}}{\partial\, \text{in}}$ is the derivative of the node's output with respect to its input. It depends only on the local computation at that node, and can be computed during the forward pass from the stored input values.
    \item The \textbf{global gradient} $\frac{\partial \ell}{\partial\, \text{in}}$ is the derivative of the loss with respect to that node's input. It accumulates contributions from all paths leading from that node to the loss.
\end{itemize}

The backpropagation update at each node is:
\begin{equation}
\frac{\partial \ell}{\partial\, \text{in}} = \frac{\partial \ell}{\partial\, \text{out}} \cdot \frac{\partial\, \text{out}}{\partial\, \text{in}}.
\end{equation}
In words: multiply the incoming global gradient (flowing in from the right) by the local gradient to obtain the outgoing global gradient (flowing to the left). This is applied at every node, proceeding right to left through the graph.

% - - - - - - - - - - - - - - - - - --
\subsection{Worked Examples}
\slideref{Session 4: Backprop Example: Linear Regression (Symbolic), (Numerical), Logistic Regression}

\subsubsection{Linear Regression --- Symbolic}

We work through the backpropagation pass for linear regression using the four-step decomposition. The forward pass evaluates:
\begin{align}
(1)\ & h = \theta_1 x, \quad (2)\ \yhat = \theta_0 + h, \quad (3)\ z = \yhat - y, \quad (4)\ \ell = z^2.
\end{align}
The local gradients are computed from these stored values:
\begin{align}
(1)\ & \partial h/\partial \theta_1 = x, \quad
(2)\ \partial \yhat/\partial \theta_0 = 1,\ \partial \yhat/\partial h = 1, \quad
(3)\ \partial z/\partial \yhat = 1, \quad
(4)\ \partial \ell/\partial z = 2z.
\end{align}
The backward pass proceeds from right to left, accumulating global gradients:
\begin{align}
(1)\ & \partial \ell/\partial z = 1 \cdot 2z = 2z \\
(2)\ & \partial \ell/\partial \yhat = 2z \cdot 1 = 2z \\
(3)\ & \partial \ell/\partial h = 2z \cdot 1 = 2z, \qquad \partial \ell/\partial \theta_0 = 2z \cdot 1 = 2z \\
(4)\ & \partial \ell/\partial \theta_1 = 2z \cdot x = 2zx.
\end{align}
The gradients with respect to the parameters $\theta_0$ and $\theta_1$ are what gradient descent requires.

\subsubsection{Linear Regression --- Numerical}

With concrete values $x=2$, $y=5$, $\theta_0=1$, $\theta_1=3$, the forward pass gives:
\begin{equation}
h = 6, \quad \yhat = 7, \quad z = 2, \quad \ell = 4.
\end{equation}
The local gradients evaluate to:
\begin{equation}
\partial h/\partial \theta_1 = 2, \quad \partial \yhat/\partial \theta_0 = \partial \yhat/\partial h = 1, \quad \partial z/\partial \yhat = 1, \quad \partial \ell/\partial z = 4.
\end{equation}
The backward pass gives $\partial\ell/\partial\theta_0 = 4$ and $\partial\ell/\partial\theta_1 = 8$, which are the gradients used to update the parameters.

\subsubsection{Logistic Regression}

The same graph structure applies to logistic regression with sigmoid activation and binary cross-entropy loss:
\begin{align}
(1)\ & h = \theta_0 + \theta_1 x, \quad
(2)\ \yhat = \sigma(h) = \frac{e^h}{1+e^h}, \quad
(3)\ \ell = -y\log\yhat - (1-y)\log(1-\yhat).
\end{align}
The local gradients are:
\begin{equation}
\partial h/\partial \theta_0 = 1, \quad \partial h/\partial \theta_1 = x, \quad
\partial \yhat/\partial h = \sigma(h)(1-\sigma(h)), \quad
\partial \ell/\partial \yhat = \frac{\yhat - y}{\yhat(1-\yhat)}.
\end{equation}
Backpropagation proceeds identically: right to left, multiplying incoming global by local at each node. The graph structure is unchanged; only the node functions differ. This is the key insight: backpropagation is a general algorithm that operates on the graph, independent of what functions the nodes implement.

% - - - - - - - - - - - - - - - - - --
\subsection{Multiple Neurons and Gradient Accumulation}
\slideref{Session 4: Backprop: Multiple Neurons in a Layer}

When a node's output feeds into multiple downstream nodes, the chain rule generalisation applies: the global gradient at that node is the \emph{sum} of contributions from each downstream path. Concretely, if node $h$ feeds into nodes $u$ and $v$:
\begin{equation}
\frac{\partial \ell}{\partial h} = \frac{\partial \ell}{\partial u} \cdot \frac{\partial u}{\partial h} + \frac{\partial \ell}{\partial v} \cdot \frac{\partial v}{\partial h}.
\end{equation}
In implementation, this is handled by initialising gradients to zero and accumulating with \texttt{+=}: each downstream node adds its contribution to the gradient of the shared upstream node. This is why the pseudocode uses \texttt{h.grad += ...} rather than \texttt{h.grad = ...}.

% - - - - - - - - - - - - - - - - - --
\subsection{Technical Implementation and Vectorisation}
\slideref{Session 4: Technical Implementation, From For Loops to Vectorization}

Each node in the computation graph stores two quantities: \texttt{*.val} (the forward value) and \texttt{*.grad} (the accumulated gradient). Node types differ in what they store and compute:

\begin{itemize}
    \item \textbf{Input nodes} store \texttt{*.val} only --- they have no gradient.
    \item \textbf{Parameter nodes} store both \texttt{*.val} and \texttt{*.grad} --- these gradients are used to update the parameters.
    \item \textbf{Compute nodes} store both \texttt{*.val} and \texttt{*.grad}, and implement both a forward function and a local gradient function.
\end{itemize}

The forward pass initialises all \texttt{*.val} by evaluating nodes from left to right. The backward pass initialises all \texttt{*.grad} to zero, sets \texttt{l.grad = 1} (the seed), and then accumulates gradients from right to left using \texttt{grad += incoming\_global * local\_grad}.

In practice, the scalar for-loop implementation is far too slow for large networks. The key optimisation is \textbf{vectorisation}: replace loops over individual neurons with matrix operations. For a single layer:
\begin{align}
&\text{Forward:} \quad \texttt{z.val = W.val @ x.val + b.val}, \quad \texttt{a.val = act(z.val)} \\
&\text{Backward:} \quad \texttt{z.grad = a.grad * dact(z.val)}, \quad \texttt{W.grad = z.grad[:, None] * x.val[None, :]}.
\end{align}
The vectorised version computes the same quantities as the for-loop, but exploits the parallelism of modern hardware. On a GPU, this can be $\sim\!100\times$ faster. This is precisely the computation that \texttt{loss.backward()} performs in PyTorch --- automatically, for any computation graph constructed from differentiable operations.
